Gas exchange in large rivers controlled by largest turbulent eddies: implications for remotely sensing gas exchange via SWOT

### 2021-07-28 manuscript_v4.Rmd
pete notes and ms updates

##1 Introduction
Natural systems play a fundamental role in the budgeting and accounting of the global carbon cycle under climate change. Since the publication of @colePlumbingGlobalCarbon2007a, the global river network is recognized to emit substantial amounts of carbon in addition to exporting to the oceans. Current estimates of total carbon dioxide evasion ($FCO_2$) to the atmosphere from the global river network vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a] **maybe add Shaoda here**, with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Despite its incredibly small percentage of the global land surface [0.47%- @raymondGlobalCarbonDioxide2013a], this flux is on par with the total oceanic $CO_2$ uptake rate [@gruberOceanicSinkAnthropogenic2019; @horgbyUnexpectedLargeEvasion2019] and the global forest carbon uptake rate [@panLargePersistentCarbon2011]. It is still relatively poorly constrained but is clearly a critical component of the global carbon cycle. Equation 1 represents this flux given the gas concentration gradient between the water $CO_{2_{water}}$ and the air $CO_{2_{air}}$ and the gas exchange velocity *k*.

@@

The structure of equation 1 necessitates that calculations of this flux are highly sensitive to one's measurements/estimates of *k* [@hallGasExchangeStreams2020]. Even more broadly, fluvial *k* is of significant importance to aquatic ecologists modeling stream metabolism [e.g. @bernhardtMetabolicRegimesFlowing2018] and water quality engineers modeling river responses to pollutant loadings [among other things- @chapraSurfaceWaterQualityModeling2008]. Decades of work have focused on elucidating the physical mechanisms behind *k* in oceans and lakes [@wanninkhofAdvancesQuantifyingAirSea2009] and, to a lesser extent, fluvial gas exchange [@hallGasExchangeStreams2020]. Given that *k* should scale with turbulence in a turbulent flow, a generalized model for *k* (equation 2) relies on the characteristic turbulence velocity scale $v_K$, which itself is a function of the characteristic length and time scales [@katulStructureFunctionModel2018]. Extensive field and labratory experiments have confirmed the validity of this model when $v_K$ is set to the Kolmogorov velocity scale [e.g. @lamontEddyCellModel1970; @zappaEnvironmentalTurbulentMixing2007; @lorkeUnifiedScalingRelation2006; @zappaVariationSurfaceTurbulence2003; @moogStreamReaerationNonuniform1999; @wangCoefficientsSmallEddy2015; @vachonRelationshipNearsurfaceTurbulence2010], where the idea is that turbulent kinetic energy (TKE) is produced by the largest, integral-scale turbulent eddies in the flow and then passed down the energy cascade to progressively smaller eddies until eddy motion is stable and TKE is dissipated as heat at the Kolmogorov-scale: the smallest eddies in a turbulent flow. Equation 2 via Kolmogorov-scale eddies can also be derived from both classic thin-film and surface renewal theories for gas exchange [@wangPhysicallyBasedScaling2021].
$$\mathbf{(2)} k\sim Sc^{-n}v_K$$

While this model works reasonably well in non-fluvial environments, there is considerable uncertainity in how it applies to fluvial systems. Specifically, the influence of the riverbed and its large roughness elements in small streams (i.e. bedforms, logjams, sediment out of suspension) complicates scaling fluvial *k* and might lead to substantially higher *k* in streams with whitewater [e.g. @hallAirWaterOxygen2012; @ulsethDistinctAirWater2019a]. Conversely, even less attention has been paid to the other side of the stream-to-river continumn, which is often hypothesized, in relation to *k*, as a hybrid condition of both fluvial and non-fluvial (i.e. lentic or estuarine) hydraulic properties. To date, the handful of existing field studies of large-river *k* have suggested that *k* begins to be notably influenced by wind dynamics, though little else is well-established [@alinPhysicalControlsCarbon2011; @beaulieuControlsGasTransfer2012a; @wangPhysicallyBasedScaling2021]. These mechanistic uncertainties are additionally limited by a large dearth of field-measured fluvial *k*. @wangPhysicallyBasedScaling2021 attempted to address this by simulating *k* in 35 rivers using stream metabolism modeling [@applingOvercomingEquifinalityLeveraging2018] and in situ dissolved oxygen (DO) datasets, finding that equation 2 is valid in their simulated rivers and that *k*~streamflow relationships are different in large and small rivers. However, they did not directly compare against *k* data in extremely small systems [@ulsethDistinctAirWater2019a] and stopped short of parsing out any hydraulic explanations for why this happens.

Further, these mechanistic uncertainties are then propogated through upscaling workflows when biogeochemists predict *k* across thousands of rivers [e.g @borgesGloballySignificantGreenhousegas2015a; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a] via equation 1 coupled with hydraulic geometry [HG: the scaling relationships between streamflow and river channel hydraulics- @leopoldHydraulicGeometryStream1953a], and it is currently not well understood how sensitive global estimates of fluvial gas evasion are to the specific HG model that is employed by the worker. Further, these approaches rely on either in situ discharge records or modeled streamflow which introduces additional uncertainities. This is all exacerbated in ungauged basins that cover large areas, especially in the carbon-rich Arctic inland waters, where little in situ information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

@@

TKE and gas exchange theory (section 2) are explored using two datasets. First, we use a previously published compilation of field measurements that were originally made to calibrate United States Geological Survey (USGS) streamgauge rating curves and gathered by @brinkerhoffReconcilingAtaStationAtManyStations2019. That dataset contains over 530,000 unique measurements of river channel velocity, width, depth, and discharge from across the continental United States. This is to our knowledge the largest collection of field measurements of river channel hydraulics and serves as the best dataset to calculate water surface TKE budgets. Second, we use the dataset from @ulsethDistinctAirWater2019a who measured and/or collected from the literature over 700 measurements of stream hydraulics and $k_{600}$ and is, to our knowledge, the largest collection of field-measured river and stream $k_{600}$. This is also the dataset used to validate our *k* scaling models.

BIKER validation (section 4) was performed on 47 SWOT-simulated rivers. Because SWOT has yet to launch, it is standard practice to benchmark SWOT-related algorithms on "SWOT-like" data. These simulated rivers are simply river-reach-averaged hydraulic model outputs where the water surface heights and widths are labelled as RS observations and are used as the sole inputs to BIKER. Here, we use 47/51 rivers collected by @frassonExploringFactorsControlling2021 and @durandIntercomparisonRemoteSensing2016b. These are the two benchmarking studies that have explored RSQ algorithm performance for the SWOT mission. The approximate locations of these rivers are plotted in Figure 1, and please consult both of those papers for all of the hydraulic model specifications. Ultimately, the 47 rivers are spread across the United States, France, Italy, the United Kingdom, and Bangladesh. We omit three models from @durandIntercomparisonRemoteSensing2016b because they lack enough hydraulic information to calculate $U_\star$, which is necessary for algorithm validation. These are both Saint Lawrence River models and the Tanana River model.

@@

##3 The TKE budget in large rivers and implications for scaling fluvial gas exchange velocity
Turbulence is fundamental to fluvial geomorphology and open-channel flows and has been the subject of extensive research across scientific and engineering disciplines for over a century. Turbulent kinetic energy (TKE) is the energy generated and dissipated by the eddies created from the chaotic nature of turbulent flow. @nakagawaTurbulenceOpenChannel1993 outline the standard energy budget for TKE, at some depth *h*, for a 2-dimensional open-channel flow (equation 3). *G* is the TKE produced, $\epsilon$ is the TKE dissipated, $T_d$ is TKE transported from the riverbed via turbulent diffusion, and $V_d$ is TKE transported via viscous diffusion. See Appendix A for all variable descriptions and notation.
$$\mathbf{(3)} G+T_d+V_d=\epsilon$$

$\epsilon$ at the free surface is frequently described as $\epsilon(H)=\frac{U_*^3}{H}$. However in rivers, which exhibit additional depth-scale shear due to downslope flow over 'macroroughness elements' (i.e. channel geometry, bars, meanders, etc.), energy dissipation occurs much more evenly over the flow depth than this parameterization allows for. It has been shown that, for rivers at least, a more appropriate dissipation model considers stream power (per unit weight water: $\epsilon_D = gSU$), which represents the energy dissipated by the total frictional resistance of the river reach's geometry [@moogStreamReaerationNonuniform1999]. This is often called 'energy dissipation via form-drag' in the literature, and is generally used to upscale *k* in rivers and streams [i.e. @raymondScalingGasTransfer2012a; @ulsethDistinctAirWater2019a]. Both @raymondScalingGasTransfer2012a and @moogStreamReaerationNonuniform1999 confirmed that $\epsilon_D$ is far greater than $\epsilon$ in open-channel flows, therefore generating the majority of fluvial TKE dissipation.

@@

We test this hypothesis using the data from @brinkerhoffReconcilingAtaStationAtManyStations2019 and @ulsethDistinctAirWater2019a by plotting boxplots of $\frac{R_h}{H}$ ratios for over 531,000 different measurements of river hydraulics versus $\epsilon_D - G$, which is equivalent to $T_d$ in equation 3 assuming that $V_d$ is neglibile away from the wall [@nakagawaTurbulenceOpenChannel1993]. *G* is calculated as $\frac{UU_*^2}{H}$. In order to calculate when $R_h=H$, we round all ratios to 3 significant figures and set all ratios greater than or equal to 0.995 as $R_h=H$. If our hypothesis holds, then $\epsilon_D - G$ should be functionally zero only when the 'large river' condition is met.

![Figure 2: Hydraulic radius/mean depth ratios (Rh/H) versus stream power per unit weight water - TKE production (eD-G at the free surface) for our two datasets of field-measured river channel hydraulics. eD-G is functionally zero only when a river is sufficiently large. This indicates that a local equilibrium of TKE is acheived and energy generated by large-scale eddies is nearly equal to energy dissipated by the smallest scale eddies. In this figure, Rh/H bins contain between 573 and 138,307 measurements.](cache/k600_theory/turbulence.jpg)

Figure 2, which plots this test, shows that $\epsilon_D-G$, or equivalently $T_d$, is nearly two orders of magnitude less when $R_h=H$ than it is when $\frac{R_h}{H}$ is between 90-99% (median $T_d$ across both datasets of $10^{-5.39}$ versus $10^{-3.83} \frac{J}{kg*s}$, respectively). When the $\frac{R_h}{H}$ ratio is very small (0-9%), this rises to a median value of $10^{-2.56} \frac{J}{kg*s}$, or nearly three orders of magnitude. We deem the $\epsilon_D - G$ values when $R_h=H$ (which are generally $10^{-5}$ to $10^{-8} \frac{J}{kg*s}$) to be functionally zero. This means that a local equilibrium of free-surface TKE is achieved, where energy produced via the largest scale (turbulent) eddies is equivalent to energy dissipated via the smallest scale (Kolmorogov) eddies. This idea of a local equilibrium is certainly not new, however this is to our knowledge the first confirmation that it happens naturally in most large rivers and that is arises algebraically from the TKE budget in open-channel flows. From here on we refer to the $R_h=H$ river scenario as the 'large river' condition.

This finding as a significant impact on how we model *k* in large rivers. Because the local equilibrium acheived during the 'large river condition' necessitates that $\epsilon$ is equivalent to *G* (Figure 2), we can presumably set $v_K$ equal to the largest ('integral-scale') eddies in the system, which are much more easily remotely-sensible than the Kolmogorov velocity scale. Otherwise, *k* should continue to scale only using the Kolmogorov-scale (which is consistent across all turbulent flows by its definition).

To test this hypothesis, we fit the following three models to the @ulsethDistinctAirWater2019a dataset of *k* measurements normalized to a Schmidt number of 600: $k_{600}=\beta U_*$, $k_{600}=\alpha \epsilon^{1/4}$, and $k_{600}=a (\epsilon_D)^b$ under two scenarios: $R_h=H$ and $R_h =/= H$. Note that because all data are normalized to a Schmidt number of 600, both the Schmidt numbers and viscosity terms are constant and are implicit in the statistical coefficients. The model structures (and the fit parameters) are detailed in Table 1. The third model tested (rows 5 and 6 of Table 1), was fit to see which slope coefficient is retrieved empirically by fitting the generalized scaling function to the data. Therfore, the $r2$ for these models is reflective of the log-transformed $k_{600}$ (unlike the other two models). This was the approach used by @ulsethDistinctAirWater2019a, who observed slopes steeper than 0.25 in both regimes of their breakpointed regression model.

@@

The results of this test are presented as Figure 3 and Table 1, where our hypothesis is confirmed: both the $U_*$ and $\epsilon^{1/4}$ models explain ~70% of variance in *k* under the 'large river' condition while the $U_*$ model does not work when this condition is not met ($r^2$:`r round(models_ustar[2,]$r2,2)` versus $r^2$=`r round(models_ustar[1,]$r2,2)`). In fact, Figures 3a and 3b are nearly visually identical in their scaling dynamics, while Figures 3d and 3e are very different, lending further evidence that these two scaling models are synonmous only when the 'large river' condition is met. This is in line with @wangPhysicallyBasedScaling2021, who found similar predictive performance and parameter values for these two models on 35 simulated rivers which they assumed large enough that $R_h \approx H$. Further, the 1/4 slope parameter is statistically recovered in the 'large river' condition (row 3, Table 1: slope = `r round(models_ulseth[2,]$slope,2)`). We note that the large residuals in Figure 3c are a byproduct of the model existing in log-space and that panels 3b and 3c are functionally identical (see Table 1 for their nearly identical intercept and slope parameters). The top row of Figure 3 confirms that under the 'large river condition' *k* can be scaled using either integral or Kolmogorov turbulence scales and that they are synonymous. This opens the door for remotely sensing *k* using almost entirely remotely-sensed parameters. This is elaborated on in section 4.

@@

The 1/4 exponent is clearly not valid when $R_h=/=H$ (Figure 3e). However, the generalized form of this model explains `r 100*round(models_ulseth[1,]$r2,2)`% of the variation in log-transformed *k* and was deemed a strong scaling model for fluvial *k* when rivers do not meet the 'large river' condition. That model yields a slope of `r round(models_ulseth[1,]$slope,2)` which corroborates other models with slopes steeper than 1/4 in rivers [@wangPhysicallyBasedScaling2021; @ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a]. Further theoretical modeling is needed to parse out why fluvial *k* in systems that do not meet the 'large river condition' generally scale at a steeper rate than in non-fluvial systems, and how this relates to the breakpointed model proposed by @ulsethDistinctAirWater2019a. The TKE budget in equation 3 suggests that it is a function of turbulent diffusion of additional TKE from the bed to the free surface.

##4 Exploiting the 'large river' model to remotely sense gas exchange velocity via the BIKER algorithm
We have shown that scaling $k_{600}$ by either integral or Kolmogorov turbulence scales explains ~70% of the variation in $k_{600}$ in large rivers and that their scaling dynamics are virtually identical. The integral-scale model relies solely on the shear velocity $U_*$. Because of the structure of $U_*$ (Appendix A), we can easily reduce the 'large river' $k_{600}$ model (Figure 3a, row 1 Table 1) to an equation consisting solely of remotely-sensible river hydraulics and an inferred estimate of the channel area. By scaling *k* this way, we significantly reduce potential equifinality issues by having just two unknowns: $k_{600}$ and the median channel area $A_0$. Equifinality refers to an under-constrained mathematical system that has essentially infinite parameter combinations that can produce the same result: there are in essence more unknowns than equations [@garamboisInferenceEffectiveRiver2015]. This problem is experienced by both SWOT RSQ algorithms and existing in situ techniques that concurrently solve for *k* and stream metabolism [@applingOvercomingEquifinalityLeveraging2018; @graceFastProcessingDiel2015; @holtgrieveSimultaneousQuantificationAquatic2010]. In both of these domains, the other parameters are quite difficult to estimate (bed roughness and stream metabolism, respectively) while for BIKER, median channel area is relatively easy to approximate from the SWOT-observable river width [@brinkerhoffConstrainingRemoteRiver2020a].

That said, this all assumes that the 'large river' condition is met. Because SWOT will be limited to the widest rivers on Earth (minimum average width of 100m), the $R_h=H$ assumption is nearly always valid. First, we verified that this is true using both the @brinkerhoffReconcilingAtaStationAtManyStations2019 and @ulsethDistinctAirWater2019a datasets (Figure S3), finding that the vast majority of rivers with a width of at least 100m exhibit this $R_h=H$ behavior. Thus, our theoretical findings in sections 2 and 3 should hold in nearly all rivers that SWOT will observe and we should be able to remotely estimate $k_{600}$ via SWOT. BIKER is informed by the @hagemannBAMBayesianAMHGManning2017a algorithm for ungauged RSQ, further explored in more recent work by @brinkerhoffConstrainingRemoteRiver2020a. These papers conceptualize discharge as a Bayesian remote sensing problem, which we largely follow here to conceptualize $k_{600}$ as a Bayesian remote sensing problem that can be solved using SWOT data.

@@

The heart of BIKER is its reformulation of the $k_{600}$ model as a Bayesian sampling model that is conditional only on the data that SWOT will provide. To do this, $k_{600}$ first needs to be written as a function of SWOT-observables *W* and $S_h$. This algebra is carried out using the model parameter from row 1 of Table 1 and yields equation 6, where we follow our earlier assumption that $R_h = H =\frac{A}{W}$ where *A* is wetted channel area. *A* is further split into the SWOT-observable portion *dA* and the unobservable portion $A_0$ following @durandEstimatingReachaveragedDischarge2014a. *dA* is estimated assuming a rectangular river channel (usually valid when $R_h=H$) so that $dA \approx \delta H*W$. Thus, there are no reliances on bed roughness, flow velocity, or any other terms that are difficult to infer from river width in BIKER.

>>>>

##1 Introduction
Natural systems play a fundamental role in the budgeting and accounting of the global carbon cycle under climate change. Since the publication of @colePlumbingGlobalCarbon2007a, the global river network is recognized to emit substantial amounts of carbon to the atmosphere, in addition to exporting it to the oceans. Current estimates of total carbon dioxide evasion ($FCO_2$) to the atmosphere from the global river network vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a] **maybe add Shaoda here**, with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Despite its incredibly small percentage of the global land surface [0.47%- @raymondGlobalCarbonDioxide2013a], this flux is on par with the total oceanic $CO_2$ uptake rate [@gruberOceanicSinkAnthropogenic2019; @horgbyUnexpectedLargeEvasion2019] and the global forest carbon uptake rate [@panLargePersistentCarbon2011]. It is still relatively poorly constrained but is clearly a critical component of the global carbon cycle. Equation 1 represents this flux given the gas concentration gradient between the water $CO_{2_{water}}$ and the air $CO_{2_{air}}$ and the gas exchange velocity *k*.

@@

The structure of equation 1 necessitates that calculations of this flux are highly sensitive to one's measurements/estimates of *k* [@hallGasExchangeStreams2020]. Even more broadly, fluvial *k* for various dissolved gases are of significant importance to aquatic ecologists modeling stream metabolism [e.g. @bernhardtMetabolicRegimesFlowing2018] and water quality engineers modeling river responses to pollutant loadings [among other things- @chapraSurfaceWaterQualityModeling2008]. Decades of work have focused on elucidating the physical mechanisms behind *k* in oceans and lakes [@wanninkhofAdvancesQuantifyingAirSea2009] and, to a lesser extent, fluvial gas exchange [@hallGasExchangeStreams2020]. Given that *k* should scale with turbulence in a turbulent flow [@hallGasExchangeStreams2020], a generalized model for *k* (equation 2) relies on the characteristic turbulence velocity scale $v_K$, which itself is a function of the characteristic length and time scales [@katulStructureFunctionModel2018]. Extensive field and labratory experiments have confirmed the validity of this model when $v_K$ is set to the Kolmogorov velocity scale [e.g. @lamontEddyCellModel1970; @zappaEnvironmentalTurbulentMixing2007; @lorkeUnifiedScalingRelation2006; @zappaVariationSurfaceTurbulence2003; @moogStreamReaerationNonuniform1999; @wangCoefficientsSmallEddy2015; @vachonRelationshipNearsurfaceTurbulence2010], where the idea is that turbulent kinetic energy (TKE) is produced by the largest, integral-scale turbulent eddies in the flow and then passed down the energy cascade to progressively smaller eddies until eddy motion is stable and TKE is dissipated as heat at the Kolmogorov-scale: the smallest eddies in a turbulent flow. Equation 2 via Kolmogorov-scale eddies can also be derived from both classic thin-film and surface renewal theories for gas exchange [@wangPhysicallyBasedScaling2021].
$$\mathbf{(2)} k\sim Sc^{-n}v_K$$

While this model works reasonably well in non-fluvial environments, there is considerable uncertainity in how it applies to fluvial systems. Specifically, the influence of the riverbed and its large roughness elements in small streams (i.e. bedforms, logjams, sediment out of suspension) complicates scaling fluvial *k* and might lead to substantially higher *k* in streams with whitewater [e.g. @hallAirWaterOxygen2012; @ulsethDistinctAirWater2019a]. Less attention has been paid to the other side of the stream-to-river continumn, which is often conceptualized in relation to *k* as a hybrid condition of both fluvial and non-fluvial hydraulic properties. To date, the handful of existing field studies of large-river *k* have suggested that *k* begins to be notably influenced by wind dynamics, though little else is well-established [@alinPhysicalControlsCarbon2011; @beaulieuControlsGasTransfer2012a; @wangPhysicallyBasedScaling2021]. These mechanistic uncertainties are additionally limited by a large dearth of field-measured fluvial *k*. @wangPhysicallyBasedScaling2021 attempted to address this by simulating *k* in 35 rivers using stream metabolism modeling [@applingOvercomingEquifinalityLeveraging2018] and in situ dissolved oxygen (DO) datasets, finding that equation 2 is valid in their simulated rivers and that *k*~streamflow relationships breakdown in large rivers. However, they did not directly compare against *k* data in extremely small systems [@ulsethDistinctAirWater2019a] and stopped short of parsing out hydraulic explanations for why this happens.

Further, these mechanistic uncertainties are then propogated through upscaling workflows when biogeochemists predict *k* across thousands of rivers [e.g @borgesGloballySignificantGreenhousegas2015a; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a] via equation 1 coupled with hydraulic geometry [HG: the scaling relationships between streamflow and river channel hydraulics- @leopoldHydraulicGeometryStream1953a]. It is currently not well understood how sensitive global estimates of fluvial gas evasion are to the specific HG model that is employed by the worker. Further, these approaches rely on either in situ discharge records or modeled streamflow which introduces additional uncertainities. This is all exacerbated in ungauged basins that cover large areas, especially in the carbon-rich Arctic inland waters, where little in situ information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

@@

TKE and gas exchange theory (section 2) are explored using two datasets. First, we use a previously published compilation of field measurements that were originally made to calibrate United States Geological Survey (USGS) streamgauge rating curves and gathered by @brinkerhoffReconcilingAtaStationAtManyStations2019. That dataset contains over 530,000 unique measurements of river channel velocity, width, depth, and discharge from across the continental United States. This is to our knowledge the largest collection of field measurements of river channel hydraulics. Second, we use the dataset from @ulsethDistinctAirWater2019a who measured and/or collected from the literature over 700 measurements of stream hydraulics and $k_{600}$ and is, to our knowledge, the largest collection of field-measured river and stream $k_{600}$. This is also the dataset used to validate our *k* scaling models.

BIKER validation (section 4) was performed on 47 SWOT-simulated rivers. Because SWOT has yet to launch, it is standard practice to benchmark SWOT-related algorithms on "SWOT-like" data. These simulated rivers are simply river-reach-averaged hydraulic model outputs where the water surface heights and widths are labelled as RS observations and are used as the sole inputs to BIKER. Here, we use 47/51 rivers collected by @frassonExploringFactorsControlling2021 and @durandIntercomparisonRemoteSensing2016b. These are the two benchmarking studies that have explored RSQ algorithm performance for the SWOT mission. The approximate locations of these rivers are plotted in Figure 1, and please consult both of those papers for all of the hydraulic model specifications. Ultimately, the 47 rivers are spread across the United States, France, Italy, the United Kingdom, and Bangladesh. We omit three models from @durandIntercomparisonRemoteSensing2016b because they lack enough hydraulic information to calculate the shear velocity $U_\star$, which is necessary for algorithm validation (see Appendix A for all variable descriptions and notation used in this study). These are two models for the Saint Lawrence River and one for the Tanana River.

@@

##3 The TKE budget in large rivers and implications for scaling fluvial gas exchange velocity
Turbulence is fundamental to fluvial geomorphology and open-channel flows and has been the subject of extensive research across scientific and engineering disciplines for over a century. Turbulent kinetic energy (TKE) is the energy generated and dissipated by the eddies created from the chaotic nature of turbulent flow. @nakagawaTurbulenceOpenChannel1993 outline the standard energy budget for TKE, at some depth *h*, for a 2-dimensional open-channel flow (equation 3). *G* is the TKE produced, $\epsilon$ is the TKE dissipated, $T_d$ is TKE transported from the riverbed via turbulent diffusion, and $V_d$ is TKE transported via viscous diffusion.
$$\mathbf{(3)} G+T_d+V_d=\epsilon$$

$\epsilon$ at the free surface is frequently described as $\epsilon(H)=\frac{U_*^3}{H}$ where *H* is the flow depth. However in rivers, which exhibit additional depth-scale shear due to downslope flow against the channel geometry, bars, and meanders, energy dissipation occurs much more evenly over the flow depth than this parameterization allows for. It has been shown that, for rivers at least, a more appropriate dissipation model considers stream power (per unit weight water: $\epsilon_D = gSU$), which represents the energy dissipated by the total frictional resistance of the river reach's geometry [@moogStreamReaerationNonuniform1999]. This is often called 'energy dissipation via form-drag' in the literature, and is generally used to upscale *k* in rivers and streams [i.e. @raymondScalingGasTransfer2012a; @ulsethDistinctAirWater2019a]. Both @raymondScalingGasTransfer2012a and @moogStreamReaerationNonuniform1999 confirmed that $\epsilon_D$ is far greater than $\epsilon$ in open-channel flows, therefore generating the majority of fluvial TKE dissipation.

@@

We test this hypothesis using the data from @brinkerhoffReconcilingAtaStationAtManyStations2019 and @ulsethDistinctAirWater2019a by plotting boxplots of $\frac{R_h}{H}$ ratios for over 531,000 different measurements of river hydraulics versus $\epsilon_D - G$, which is equivalent to $T_d$ in equation 3 assuming that $V_d$ is neglibile away from the wall [@nakagawaTurbulenceOpenChannel1993]. *G* is calculated as $\frac{UU_*^2}{H}$. In order to determine when $R_h=H$, we round all ratios to 3 significant figures and set all ratios greater than or equal to 0.995 as $R_h=H$. If our hypothesis holds, then $\epsilon_D - G$ should be functionally zero only when the 'large river' condition is met.

![Figure 2: Rh/H versus eD-G at the free surface for our two datasets of field-measured river channel hydraulics (see Appendix A for variable descriptions). eD-G is functionally zero only when a river is sufficiently large. This indicates that a local equilibrium of TKE is acheived and energy generated by large-scale eddies is nearly equal to energy dissipated by the smallest scale eddies. In this figure, Rh/H bins contain between 573 and 138,307 measurements.](cache/k600_theory/turbulence.jpg)

Figure 2, which plots this test, shows that $\epsilon_D-G$ is nearly two orders of magnitude less when $R_h=H$ than it is when $\frac{R_h}{H}$ is between 90-99% (median value across both datasets of $10^{-5.39}$ versus $10^{-3.83} \frac{J}{kg*s}$, respectively). When the $\frac{R_h}{H}$ ratio is very small (0-9%), this rises to a median value of $10^{-2.56} \frac{J}{kg*s}$, or nearly three orders of magnitude larger. The $\epsilon_D - G$ values when $R_h=H$ (which are generally $10^{-5}$ to $10^{-8} \frac{J}{kg*s}$) are functionally zero. This means that a local equilibrium of free-surface TKE is achieved, where energy produced via the largest scale (turbulent) eddies is equivalent to energy dissipated via the smallest scale (Kolmorogov) eddies. This idea of a local equilibrium is certainly not new, however this is to our knowledge the first confirmation that it happens naturally in most large rivers and that is arises algebraically from the TKE budget in open-channel flows. From here on we refer to the $R_h=H$ river scenario as the 'large river' condition.

This finding has a significant impact on how we model *k* in large rivers. Because the local equilibrium acheived during the 'large river condition' necessitates that $\epsilon$ is equivalent to *G* (Figure 2), we can presumably set $v_K$ in equation 2 equal to the largest ('integral-scale') eddies in the system, which are much more easily remotely-sensible than the Kolmogorov velocity scale. Otherwise, *k* should continue to scale only using the Kolmogorov-scale (which is consistent across all turbulent flows by its definition).

To test this hypothesis, we fit the following three models to the @ulsethDistinctAirWater2019a dataset of *k* measurements normalized to a Schmidt number of 600: $k_{600}=\beta U_*$, $k_{600}=\alpha \epsilon^{1/4}$, and $k_{600}=a (\epsilon_D)^b$ under two scenarios: $R_h=H$ and $R_h =/= H$. Note that because all data are normalized to a Schmidt number of 600, both the Schmidt numbers and viscosity terms are constant and are implicit in the statistical coefficients. The model structures (and the fit parameters) are detailed in Table 1. The third model tested (rows 5 and 6 of Table 1), was fit to see which slope coefficient is retrieved empirically by fitting the generalized scaling function to the data. Therfore, the $r2$ for these models is reflective of the log-transformed $k_{600}$ (unlike the other two models). This was the approach used by @ulsethDistinctAirWater2019a, who observed slopes steeper than 1/4 in both regimes of their breakpointed regression model.

@@

The results of this test are presented as Figure 3 and Table 1, where our hypothesis is confirmed: both the $U_*$ and $\epsilon^{1/4}$ models explain ~70% of variance in *k* under the 'large river' condition while the $U_*$ model does not work when this condition is not met ($r^2$:`r round(models_ustar[2,]$r2,2)` versus $r^2$=`r round(models_ustar[1,]$r2,2)`). In fact, Figures 3a and 3b are nearly visually identical in their scaling dynamics, while Figures 3d and 3e are very different, lending further evidence that these two scaling models are synonmous only when the 'large river' condition is met. This is in line with @wangPhysicallyBasedScaling2021, who found similar predictive performance and parameter values for these two models on 35 simulated riverswhich they assumed were large enough that $R_h \approx H$. Further, the 1/4 slope parameter is statistically recovered in the 'large river' condition (row 3, Table 1: slope = `r round(models_ulseth[2,]$slope,2)`). We note that the large residuals in Figure 3c are a byproduct of the model existing in log-space and that panels 3b and 3c are functionally identical models (see Table 1 for their nearly identical intercept and slope parameters). The top row of Figure 3 confirms that under the 'large river condition' *k* can be scaled using either integral or Kolmogorov turbulence scales and that they are synonymous. This opens the door for remotely sensing *k* using almost entirely remotely-sensed parameters. This is elaborated on in section 4.

@@

The 1/4 exponent is clearly not valid when $R_h=/=H$ (Figure 3e). However, the generalized form of this model explains `r 100*round(models_ulseth[1,]$r2,2)`% of the variation in log-transformed *k* and was deemed a strong scaling model for fluvial *k* when rivers do not meet the 'large river' condition. That model yields a slope of `r round(models_ulseth[1,]$slope,2)` which corroborates other models with slopes steeper than 1/4 in rivers [@wangPhysicallyBasedScaling2021; @ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a]. Further theoretical modeling is needed to parse out why fluvial *k* in systems that do not meet the 'large river condition' generally scale at a steeper rate than in non-fluvial systems, and how this relates to the breakpointed model proposed by @ulsethDistinctAirWater2019a. That said, the TKE budget in equation 3 suggests that it is a function of turbulent diffusion of additional TKE from the bed to the free surface.

##4 Exploiting the 'large river' model to remotely sense gas exchange velocity via the BIKER algorithm
We have shown that scaling $k_{600}$ by either integral or Kolmogorov turbulence scales explains ~70% of the variation in $k_{600}$ in large rivers and that their scaling dynamics are virtually identical. The integral-scale model relies solely on the shear velocity $U_*$. Because of the structure of $U_*$ (Appendix A), we can easily reduce the 'large river' $k_{600}$ model (Figure 3a, row 1 Table 1) to an equation consisting solely of remotely-sensible river hydraulics and an intial estimate of the channel area. By scaling *k* this way, we significantly reduce potential equifinality issues by having just two unknowns: $k_{600}$ and the median channel area $A_0$. Equifinality refers to an under-constrained mathematical system that has essentially infinite parameter combinations that can produce the same result: there are in essence more unknowns than equations [@garamboisInferenceEffectiveRiver2015]. This problem is experienced by both SWOT RSQ algorithms and in situ tools that concurrently solve for *k* and stream metabolism [@applingOvercomingEquifinalityLeveraging2018; @graceFastProcessingDiel2015; @holtgrieveSimultaneousQuantificationAquatic2010]. In both of these domains, the other unknown parameters are often difficult to estimate (bed roughness and stream metabolism, respectively) while for BIKER, median channel area is relatively easy to approximate from the SWOT-observable river width [@brinkerhoffConstrainingRemoteRiver2020a].

That said, this all assumes that the 'large river' condition is met. Because SWOT will be limited to the widest rivers on Earth (minimum average width of 100m), the $R_h=H$ assumption is nearly always valid. First, we verified that this is true using both the @brinkerhoffReconcilingAtaStationAtManyStations2019 and @ulsethDistinctAirWater2019a datasets (Figure S3), finding that the vast majority of rivers with a width of at least 100m exhibit this $R_h=H$ behavior. Thus, our theoretical findings in sections 2 and 3 should hold in nearly all rivers that SWOT will observe and we should be able to remotely estimate $k_{600}$ via SWOT. Therefore, we develop the BIKER algorithm, which is informed by the @hagemannBAMBayesianAMHGManning2017a algorithm for ungauged RSQ and further explored in more recent work by @brinkerhoffConstrainingRemoteRiver2020a. These papers conceptualize discharge as a Bayesian remote sensing problem, which we largely follow here to conceptualize $k_{600}$ as a Bayesian remote sensing problem that can be solved using SWOT data.

@@

The heart of BIKER is its reformulation of the $k_{600}$ model as a Bayesian sampling model that is conditional only on the data that SWOT will provide. To do this, $k_{600}$ is first written as a function of SWOT-observables *W* and $S_h$. This algebra is carried out using the model parameter from row 1 of Table 1 and yields equation 6, where we follow our earlier assumption that $R_h = H =\frac{A}{W}$ where *A* is wetted channel area. *A* is further split into the SWOT-observable portion *dA* and the unobservable portion $A_0$ following @durandEstimatingReachaveragedDischarge2014a. *dA* is estimated assuming a rectangular river channel (usually valid when $R_h=H$) so that $dA \approx \delta H*W$. Thus, there are no reliances on bed roughness, flow velocity, or any other terms that are difficult to infer from river width.
